---
title: "Steps in mdEditor"
output: 
  # html_document:
  #   theme: readable
  #   highlight:
  #   fig_caption: true
  knitr:::html_vignette:
    toc: true
  pdf_document:
    highlight: null
    number_sections: yes
vignette: >
  %\VignetteIndexEntry{mdJSONdictio}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
<style type="text/css">
  body{
  font-size: 12pt;
}
</style>
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(width = 9999)

library(mdJSONdictio)
library(knitr)
library(bookdown)
library(kableExtra)
library(tibble)
library(rmarkdown)
library(DT)
library(png)
library(utils)
library(rjson)
```

# After using ```build.mdJSON()```

## 1. Import the mdJSON data dictionary to mdEditor
Open [mdEditor](https://go.mdeditor.org/#/dashboard) in your browser and navigate to the Import tab at the top of the page. Drag and drop the newly written mdJSON file and click "Import Data." This will generate a Dictionary record with the dictionary title you specified in the function. You may need to refresh your browser if the dictionary does not immediately appear

## 2. Fix errors in the Dictionary record
When viewing your Dictionary record, you will notice an orange exclamation mark next to the title that indicates the dictionary is incomplete and has missing or erroneous elements. Click the orange exclamation mark to view the details. These errors can be remedied by completing the Responsible Party field in the Main tab and the Entity Code Name and Definition fields in the Entities tab. For more information about these fields, refer to the [mdEditor User Manual](https://guide.mdeditor.org/reference/reference-manual.html). You will know your dictionary is free of errors when the orange triangle disappears and the edit icon next to your Dictionary record in the left-hand menu is green. If you see more than these three errors, make sure the tabular data dictionary is filled out as described in the [Tabular Data Dictionary Template] section.

## 3. Associate and maintain the Dictionary record
Once the Dictionary record is free of errors, it can be associated with the Metadata record(s) it describes (e.g. tabular dataset, shapefiles) to help create archival quality project metadata. For details of this process and other information, refer to the [mdEditor User Manual](https://guide.mdeditor.org/). Any future updates to your Dictionary record can be done manually in the mdEditor web application. 
<br />
<br />

# Before using ```build.table()```

## 1. Complete required mdEditor fields
Open [mdEditor](https://go.mdeditor.org/#/dashboard) in your browser and import your mdJSON data dictionary file as described in step 1 above. Navigate to your Dictionary record and verify the contents and check for errors. You will know your dictionary is free of errors when the orange triangle disappears and the edit icon next to your Dictionary record in the left-hand menu is green. While ```build.table()``` only requires an entity with at least one defined attribute, an error-free Dictionary record is more likely to be exported with a structure that can be processed by the function. Please note, if your Dictionary record contains two entities, they will be processed into a single tabular data dictionary.

## 2. Export the mdJSON data dictionary
Navigate to the Export tab at the top of the page. Make sure no Metadata records, Contacts or undesired Dictionaries are selected (select and deselect the checkboxes under the search bars). Select the desired dictionary and click "Export Selected" to the right of the screen. Save the file to your working space. If you are not prompted to select a save location, the file may have automatically saved to your local Downloads folder. If you intend to transform more than one Dictionary record into a table, you can select and export them together into a single mdJSON file and specify which dictionary is processed by ```build.table()``` with the parameter ```dictionary_num```.
<br />
<br />
<br />
